A Fast Unified Model for Parsing and Sentence Understanding

### 2016-03-18 writing/hard_stack_paper/paper.tex
Add RNN speed results (costs us five lines..)

\begin{figure}
\centering
\resizebox{0.85\linewidth}{!}{
\begin{tikzpicture} 
\begin{axis}[xmin=0, ymin=0, xlabel=Batch size, ylabel=Feedforward time (sec),
             xtick={0,256,512,1024,2048}, legend style={at={(1.1, 0.98)},anchor=north east}]
  \addplot table [x={Batch size}, y=CPU, col sep=tab] {runtime.csv};
  \addlegendentry{CPU \citep{irsoy2014deep}}
  \addplot table [x={Batch size}, y=GPU, col sep=tab] {runtime.csv};
  \addlegendentry{Thin-stack GPU}
\end{axis}
\end{tikzpicture}
}
\caption{Feedforward speed comparison.}
\label{fig:speed}
\end{figure}

In this section, we compare the test time speed of our SPINN-PI-NT with an equivalent TreeRNN implemented in the conventional fashion. While the full models evaluated below are implemented and trained using Theano \citep{bergstra+al:2010-scipy,Bastien-Theano-2012}, which is reasonably efficient but not perfect for our model, we wish to compare well-optimized implementations of both models here. To do this, we reimplement the feedforward of SPINN-PI-NT in C++/CUDA, and we compare that implementation with a CPU-based C++/Eigen TreeRNN implementation from \citet{irsoy2014deep}, which we modified to perform exactly the same computations as our model.\footnote{The original code for Irsoy \& Cardie's model is available at \url{https://github.com/oir/deep-recursive}. We plan to release our modified code at publication time, alongside the optimized C++/CUDA SPINN and the Theano source code for the full SPINN.} TreeRNNs like this can only operate on a single example at a time and are thus poorly suited for GPU computation.\footnote{We chose to reimplement and evaluate only the feedforward/inference pass, as inference speed is the relevant performance metric for most practical applications.}

Each model is restricted to run on sentences of 30 tokens or fewer. We fix the model dimension and the word embedding dimension at 300. We run the CPU performance test on a 2.20-GHz Intel Xeon processor with hyperthreading enabled. Our GPU model is implemented in C++ and CUDA. We test it on a NVIDIA Titan X GPU.

Figure~\ref{fig:speed} compares the sentence encoding speed of the two models using data from the Stanford Sentiment Treebank \citep{socher2013recursive}. Figure~\ref{fig:speed} shows a substantial difference in runtime that increases with batch size, as is typical of GPU computation. At the larger batch sizes that are most effective for test-time applications (e.g., 512+ examples), we observe a 10--25x speedup relative to the standard CPU implementation.

>>>>

\begin{figure}
\centering
\resizebox{0.85\linewidth}{!}{
\begin{tikzpicture} 
\begin{axis}[xmin=0, ymin=0, xlabel=Batch size, ylabel=Feedforward time (sec),
             xtick={0,256,512,1024,2048}, legend style={at={(1.1, 0.98)},anchor=north east}]
  \addplot table [x={Batch size}, y=GPU, col sep=tab] {runtime.csv};
  \addlegendentry{Thin-stack GPU}
  \addplot table [x={Batch size}, y=CPU, col sep=tab] {runtime.csv};
  \addlegendentry{CPU \citep{irsoy2014deep}}
  \addplot table [x={Batch size}, y=RNN, col sep=tab] {runtime.csv};
  \addlegendentry{RNN}
\end{axis}
\end{tikzpicture}
}
\caption{Feedforward speed comparison.}
\label{fig:speed}
\end{figure}

In this section, we compare the test time speed of our SPINN-PI-NT with an equivalent TreeRNN implemented in the conventional fashion and a standard recurrent neural network sequence model. While the full models evaluated below are implemented and trained using Theano \citep{bergstra+al:2010-scipy,Bastien-Theano-2012}, which is reasonably efficient but not perfect for our model, we wish to compare well-optimized implementations of both models here. To do this, we reimplement the feedforward of SPINN-PI-NT in C++/CUDA, and we compare that implementation with a CPU-based C++/Eigen TreeRNN implementation from \citet{irsoy2014deep}, which we modified to perform exactly the same computations as our model.\footnote{The original code for Irsoy \& Cardie's model is available at \url{https://github.com/oir/deep-recursive}. We plan to release our modified code at publication time, alongside the optimized C++/CUDA SPINN and the Theano source code for the full SPINN.} TreeRNNs like this can only operate on a single example at a time and are thus poorly suited for GPU computation.\footnote{We chose to reimplement and evaluate only the feedforward/inference pass, as inference speed is the relevant performance metric for most practical applications.} We also include speed results from a simple C++/CUDA implementation of a recurrent neural network.

Each model is restricted to run on sentences of 30 tokens or fewer. We fix the model dimension and the word embedding dimension at 300. We run the CPU performance test on a 2.20-GHz Intel Xeon processor with hyperthreading enabled. We test our thin-stack implementation and the RNN model on an NVIDIA Titan X GPU.

Figure~\ref{fig:speed} compares the sentence encoding speed of the three models on random input data. Figure~\ref{fig:speed} shows a substantial difference in runtime between the CPU and thin-stack implementations that increases with batch size. At the larger batch sizes that are most effective for test-time applications (e.g., 512+ examples), we observe a 10--25x speedup relative to the standard CPU implementation. The thin-stack implementation is consistently about 4x slower than the recurrent neural network we tested here.
